Inverse energy transfer in three-dimensional quantum vortex flows

### 2025-06-19 revision/referee-response-2.tex

    \refcomment{1) p.1, right, "helicity, which is also an inviscid invariant": During
    the reconnection in a finite temperature, the total helicity including
    mutual friction force and normal fluid velocity is conserved? Normal
    fluid is handled as a viscous flow. Please describe to readers.}

    The presence of the mutual friction force introduces an \red{additional source of} dissipation into the normal fluid. By \red{also} computing the helicity in the superfluid component, we observe that during the reconnection event
the total helicity \red{(sum of normal fluid helicity and superfluid helicity)} is not conserved. We omitted \red{the plot of the superfluid helicity}
from the \red{original} manuscript so as to \red{avoid introducing}
\red{the further concept} 
of superfluid helicity to a paper whose narrative is \red{dedicated to the
normal fluid}. \red{In the revised manuscript} we have addressed the 
referee's concern by inserting \red{(in page 4)} an additional statement 
to clarify that the total helicity during the reconnection event is 
\textbf{not} conserved. \\

    \refcomment{2) p.3, right: Two pairs of initially orthogonal vortices are
    considered in the present configuration. This paper reports that one
    pair of vortices yield a negative helicity to the normal fluid
    velocity. Please mention whether the other pair of vortices yields the
    positive helicity or not. If both of them yield the negative helicity,
    the initial setup already causes a symmetry breaking.}

    In our manuscript, we indeed only describe the results referring to one pair of reconnecting 
    vortices. We confirm that the helicity injected by the other pair is indeed positive.

%    The initial \red{vortex} configuration is set up such that the 
%superfluid velocity circulation in each half-box has the opposite sign to 
%the other half-box; \red{in this way, the superfluid velocity satisfies
%periodic boundary conditions over the entire computational 
%domain}.
%\red{It is important to appreciate that the two reconnection events do
%not occurr at the same time}. 
%In the timescale of the simulation \red{which we report in our manuscript}, 
%we only observe one pair of vortices which reconnects, as the reconnection 
%\red{of the other vortex pair occurs later, when the main impact of the first
%reconnection has vanished}.

%\red{We find that} during \red{this first reconnection}, the helicity contribution
%from the other vortex pair is negligible. \red{If we extend the
%simulation for a longer time and observe the second reconnection,
%we notice that} the second reconnection injects positive helicity 
%in the normal fluid, which then decays back to zero.

%Additionally, we computed the normal fluid helicity in each respective 
%half-box, confirming that the first reconnection injects negative
%helicity (computed only in its half-box), and, later, the second reconnection
%injects positive helicity (in its own half-box).

%To \red{better} illustrate this point, we have modified the text to 
%explicity state that \red{the two} reconnections occur 
%asynchronously, \red{that the second reconnection is required by the
%periodic boundary conditions of our computational domain}, 
%\red{and that, physically, we focus the discussion of our results
%on the first reconnection only}.
%We have also added some extra details to Fig.2 in the \red{revised}
%manuscript which highlight this \red{feature} visually. \\

### 2025-06-19 main_REV.tex
Updated referee response and revised tex files

% WE ADD REFERENCE TO OUR FIRST PAPER LATER IN THE TEXT
%Punctuated energy injection resulting from the violent nature of vortex reconnections \cite{stasiak2024quantum} pave the way for an imbalance of chirality by multi-scale energy injection.
%

@@

To study the reconnection dynamics, we consider two pairs of initially orthogonal vortices (where the corresponding vortices of each pair have opposite circulation in order to preserve periodicity along the boundaries)
at two distinct temperatures, $T=1.9K$ and $T=2.1K$. The vortex pairs are separated by the distance $D_{\ell}$; each vortex within each pair is initially at distance $d_{\ell}$ to the other vortex, such that $d_{\ell}\ll D_{\ell}$ to ensures that the dynamics in the vicinity of the reconnection is dominated by local interactions, and that the far-field contribution from the other vortex pair is negligible. \red{In our simulations, we consider only the vortex dynamics up to and shortly after the first reconnection, as the reconnection events occur asynchronously (see Fig.~\ref{fig:total-helicity}). Hence, all results reported and discussed herafter refer to the first reconnection.}.

The evolution of the vortex reconnection of a single pair is reported in Fig.~\ref{fig:visualisation}. 
The first row shows the reconnecting superfluid vortices (in green) accompanied by normal fluid structures generated by mutual friction, here displayed as enstrophy rendering $\bom(\x)^2=|\nabla\times \v_n|^2$. Such structures are the signature of the violent irreversible energy transfers in vortex reconnections 
\cite{stasiak2024quantum}. The second row shows the rendering of the local helicity $H(\x)=\v_n\cdot\bom$, where we observe a clear local helicity production, with an abrupt change of sign due to the rearrangement of the vortex topology. Remarkably, during reconnection there is a net 
sudden normal fluid helicity production, as shown in Fig.~\ref{fig:total-helicity}. We will come back to this finding later.
\begin{figure}[t]
    \centering
    \includegraphics*[width=0.48\textwidth]{helicity_normal_fluid.pdf}
\caption{Temporal evolution of the normal fluid helicity $\mathcal{H} = \int_{\mathcal{V}}H(\x)dV$ computed over the entire volume $\mathcal{V}$. For superfluid helium, it is proper to make the helicity dimensionless in terms of $\kappa^2$. \red{The 3D rendering shows the asynchronous nature of the reconnections in the initial configuration, with the same colour scheme as in Fig.~\ref{fig:visualisation}.}}
\label{fig:total-helicity}
\end{figure}

@@

A chiral imbalance occurs if the mutual friction force is helical, 
\textit{i.e.} if the ratio $|f^+|^2/|f^-|^2 \neq 1$, with $|f^\pm|^2$ the total squared norm of the mutual friction components. 
In Fig.~\ref{fig:mutual-friction-decomp}, we show the temporal evolution of $|f^+|^2/|f^-|^2$, for both temperatures. 
%
\begin{figure}[h!]
    \centering
    \includegraphics*[width=0.48\textwidth]{fmfDecompFig.pdf}
    \caption{Temporal evolution of projected mutual friction force components $f^\pm$/ Inset: temporal evolution of total helical components.}.
    \label{fig:mutual-friction-decomp}
\end{figure}
%
It is apparent
that during and after the reconnection, the mutual friction force is strongly chiral,
injecting more negative helicity than positive helicity. 
As a result, the ratio $\mathcal{H}^+/\mathcal{H}^-$ 
(reported in the inset of Fig.~\ref{fig:mutual-friction-decomp}) 
decreases significantly at reconnection and remains smaller than unity even at later times, 
indicating that the flow is persistently chiral. We conclude that the reconnection triggers indeed a chiral imbalance.
\red{
For the sake of completeness, it is worth noting that the total helicity of the flow (obtained by adding the superfluid helicity to that of the normal fluid $\mathcal{H}$) is not conserved.
}
From Fig.~\ref{fig:mutual-friction-decomp}, \red{ we consider a time $t^*$ after reconnection where the ratio $|f^+|^2/|f^-|^2$ has sufficiently decayed to a roughly 5\% difference and} we determine the non-dimensional timescale \red{$\tau = (t^*-t_R)/\tau_R$} during which the mutual friction force is chiral as a result of reconnections: $\tau \approx 0.01$ and $\tau \approx 0.005$
for $T=1.9$K and $T=2.1$K, respectively, corresponding dimensionally to $\tau \approx 0.1$s for both temperatures. 
In superfluid turbulence, the timescale between two consecutive reconnections can be smaller than $\tau$ provided that the
vortex line density $\mathcal{L}$ (length of vortices per unit volume) is larger than $10^8 \text{m}^{-2}$ \cite{stasiak2024quantum,barenghi2004},
a condition which is easily met in superfluid helium experiments \cite{roche2007,Babuin2014}.

@@

%\paragraph*{Closing remarks.---}
In conclusion, the reconnection of quantum vortices in the two-fluid regime
($T\gtrsim 1.5$K) not only injects punctuated
energy in the normal fluid \cite{stasiak2024quantum}, but also triggers in the normal fluid
a transfer of kinetic energy towards the large scales. 
%This inverse 
%energy transfer arises from the fact that the mutual friction force (injecting energy and 
%helicity in the normal fluid) is helical, as Kelvin waves develop on the vortices as a result of the reconnection itself.
%The helical character of the mutual friction produces a chiral imbalance in the normal fluid, driving this inverse transfer, as previously observed in turbulent Navier-Stokes flows
This inverse energy transfer \red{is triggered by} the helical character of the
mutual friction, \red{directly arising from the} Kelvin waves released by the reconnecting
cusp, which produces a chiral imbalance in the normal fluid, as 
previously observed in turbulent Navier-Stokes flows
\cite{biferaleInverseEnergyCascade2012a,plunianInverseCascadeEnergy2020a}.

Our findings have profound implications for the nature of turbulence in 
finite temperature superfluids, \red{where vortex reconnections play a key role}. 
In circumstances where the vortex density 
$\mathcal{L}$ is large and where the isotropy of the vortex tangle 
is broken by external forcing (for instance thermal counterflows \cite{vinen-1957a}), 
the chirality of the flow generated by the 
frequent \red{non-symmetrical} reconnections may be strong enough to induce an inverse energy 
cascade \cite{notes_recon}.
%The asymmetry of the reconnecting vortex configurations is a key element; 
%when perfectly symmetrical vortex configurations (with corresponding 
%opposite vortex line orientations) reconnect, the injected helicity 
%globally cancels out, as verified in our study if we take also into 
%account the reconnection.
The microscopic mechanism that we have described \red{may be relevant in triggering}
the inverse energy cascade which is observed numerically in  
large-scale simulations of counterflow turbulence at large heat fluxes \cite{polanco2020}, which is
indeed inherently not isotropic.
%, hence triggering non-fully reconnections. 
Our work hence motivates further detailed studies of
the role played by helicity in superfluid dynamics 
\cite{dileoni2016,galantucci2021},
moving the emphasis from few vortex systems \cite{scheeler2014} 
to fully coupled superfluid turbulence.

@@

\bibitem [{\citenamefont {Stasiak}\ \emph {et~al.}(2024)\citenamefont

@@

  {Galantucci},\ and\ \citenamefont {Krstulovic}}]{stasiak2024quantum}%

@@

  {\bibfnamefont {A.~W.}\ \bibnamefont {Baggaley}}, \bibinfo {author}

@@

  {arXiv}\ ,\ \bibinfo {pages} {2411.08942}} (\bibinfo {year}
  {2024})}\BibitemShut {NoStop}%

@@

    % \include{main_REV.bbl}
\end{document}
%
% ****** End of file apssamp.tex ******

>>>>

% WE ADD REFERENCE TO OUR FIRST PAPER LATER IN THE TEXT
%Punctuated energy injection resulting from the violent nature of vortex reconnections \cite{stasiak2025experimental} pave the way for an imbalance of chirality by multi-scale energy injection.
%

@@

To study the reconnection dynamics, we consider two pairs of initially orthogonal vortices (where the corresponding vortices of each pair have opposite circulation in order to preserve periodicity along the boundaries)
at two distinct temperatures, $T=1.9K$ and $T=2.1K$. The vortex pairs are separated by the distance $D_{\ell}$; each vortex within each pair is initially at distance $d_{\ell}$ to the other vortex, such that $d_{\ell}\ll D_{\ell}$ to ensures that the dynamics in the vicinity of the reconnection is dominated by local interactions, and that the far-field contribution from the other vortex pair is negligible. 
\red{All results reported hereafter refer only to the reconnection of one pair of vortices.} 
%\red{In our simulations, the reconnections of the two pairs of vortices occur at different times; here we consider only the vortex dynamics up to and shortly after the first reconnection (see Fig.~\ref{fig:total-helicity}). Hence, all results reported and discussed herafter refer to the first reconnection.}.

The evolution of the vortex reconnection of a single pair is reported in Fig.~\ref{fig:visualisation}. 
The first row shows the reconnecting superfluid vortices (in green) accompanied by normal fluid structures generated by mutual friction, here displayed as enstrophy rendering $\bom(\x)^2=|\nabla\times \v_n|^2$. Such structures are the signature of the violent irreversible energy transfers in vortex reconnections 
\cite{stasiak2025experimental}. The second row shows the rendering of the local helicity $H(\x)=\v_n\cdot\bom$, where we observe a clear local helicity production, with an abrupt change of sign due to the rearrangement of the vortex topology. Remarkably, during reconnection there is a net 
sudden normal fluid helicity production, as shown in Fig.~\ref{fig:total-helicity}. We will come back to this finding later.
\begin{figure}[t]
    \centering
    \includegraphics*[width=0.48\textwidth]{helicity_total.pdf}
\caption{Temporal evolution of the normal fluid helicity $\mathcal{H} = \int_{\mathcal{V}}H(\x)dV$ computed over the entire volume $\mathcal{V}$. For superfluid helium, it is proper to make the helicity dimensionless in terms of $\kappa^2$. $\tau_R = L_R/E^{1/2}_R$.
%\red{The 3D rendering shows the asynchronous nature of the reconnections in the initial configuration, with the two reconnections occurring at different times; same colour scheme as in Fig.~\ref{fig:visualisation}.}
}
\label{fig:total-helicity}
\end{figure}

@@

A chiral imbalance occurs if the mutual friction force is helical, 
\textit{i.e.} if the ratio $|f^+|^2/|f^-|^2 \neq 1$, with $|f^\pm|^2$ the total squared norm of the mutual friction components. 
In Fig.~\ref{fig:mutual-friction-decomp}, we show the temporal evolution of $|f^+|^2/|f^-|^2$, for both temperatures. 
%
\begin{figure}[h!]
    \centering
    \includegraphics*[width=0.48\textwidth]{fmfDecompFig.pdf}
    \caption{Temporal evolution of projected mutual friction force components $f^\pm$/ Inset: temporal evolution of total helical components.}.
    \label{fig:mutual-friction-decomp}
\end{figure}
%
It is apparent
that during and after the reconnection, the mutual friction force is strongly chiral,
injecting more negative helicity than positive helicity. 
As a result, the ratio $\mathcal{H}^+/\mathcal{H}^-$ 
(reported in the inset of Fig.~\ref{fig:mutual-friction-decomp}) 
decreases significantly at reconnection and remains smaller than unity even at later times, 
indicating that the flow is persistently chiral. We conclude that the reconnection triggers indeed a chiral imbalance.
\red{
For the sake of completeness, it is worth noting that the total helicity of the flow (obtained by adding the superfluid helicity to that of the normal fluid $\mathcal{H}$) is not conserved.
}
From Fig.~\ref{fig:mutual-friction-decomp}, \red{ we consider a time $t^*$ after reconnection where the ratio $|f^+|^2/|f^-|^2$ has sufficiently decayed to a roughly 5\% difference and} we determine the non-dimensional timescale \red{$\tau = (t^*-t_R)/\tau_R$} during which the mutual friction force is chiral as a result of reconnections: $\tau \approx 0.01$ and $\tau \approx 0.005$
for $T=1.9$K and $T=2.1$K, respectively, corresponding dimensionally to $\tau \approx 0.1$s for both temperatures. 
In superfluid turbulence, the timescale between two consecutive reconnections can be smaller than $\tau$ provided that the
vortex line density $\mathcal{L}$ (length of vortices per unit volume) is larger than $10^8 \text{m}^{-2}$ \cite{stasiak2025experimental,barenghi2004},
a condition which is easily met in superfluid helium experiments \cite{roche2007,Babuin2014}.

@@

%\paragraph*{Closing remarks.---}
In conclusion, the reconnection of quantum vortices in the two-fluid regime
($T\gtrsim 1.5$K) not only injects punctuated
energy in the normal fluid \cite{stasiak2025experimental}, but also triggers in the normal fluid
a transfer of kinetic energy towards the large scales. 
%This inverse 
%energy transfer arises from the fact that the mutual friction force (injecting energy and 
%helicity in the normal fluid) is helical, as Kelvin waves develop on the vortices as a result of the reconnection itself.
%The helical character of the mutual friction produces a chiral imbalance in the normal fluid, driving this inverse transfer, as previously observed in turbulent Navier-Stokes flows
This inverse energy transfer \red{is triggered by} the helical character of the
mutual friction, \red{directly arising from the} Kelvin waves released by the reconnecting
cusp, which produces a chiral imbalance in the normal fluid, as 
previously observed in turbulent Navier-Stokes flows
\cite{biferaleInverseEnergyCascade2012a,plunianInverseCascadeEnergy2020a}.

Our findings have profound implications for the nature of turbulence in 
finite temperature superfluids, \red{where vortex reconnections play a key role}. 
In circumstances where the vortex density 
$\mathcal{L}$ is large and where the isotropy of the vortex tangle 
is broken by external forcing (for instance thermal counterflows~\cite{vinen-1957a}), 
the chirality of the flow generated by the 
frequent \red{non-symmetrical} reconnections may be strong enough to induce an inverse energy 
cascade~\cite{notes_recon}.
%The asymmetry of the reconnecting vortex configurations is a key element; 
%when perfectly symmetrical vortex configurations (with corresponding 
%opposite vortex line orientations) reconnect, the injected helicity 
%globally cancels out, as verified in our study if we take also into 
%account the reconnection.
The microscopic mechanism that we have described \red{may be relevant in triggering}
the inverse energy cascade which is observed numerically in  
large-scale simulations of counterflow turbulence at large heat fluxes \cite{polanco2020}, which is
indeed inherently not isotropic.
%, hence triggering non-fully reconnections. 
Our work hence motivates further detailed studies of
the role played by helicity in superfluid dynamics 
\cite{dileoni2016,galantucci2021},
moving the emphasis from few vortex systems \cite{scheeler2014} 
to fully coupled superfluid turbulence.

@@

\bibitem [{\citenamefont {Stasiak}\ \emph {et~al.}(2025)\citenamefont

@@

  {Galantucci},\ and\ \citenamefont {Krstulovic}}]{stasiak2025experimental}%

@@

  {\bibfnamefont {A.}~\bibnamefont {Baggaley}}, \bibinfo {author}

@@

  {Proc. Nat. Acad. of Sci.}\ }\textbf {\bibinfo {volume} {122}},\ \bibinfo
  {pages} {e2426064122} (\bibinfo {year} {2025})}\BibitemShut {NoStop}%

@@

\end{document}
%
% ****** End of file apssamp.tex ******
